Monte Carlo Simulations

### 2015-03-23 mc_sim.rmd
redid analysis with 2d smooth

First, load some r packages that we will need

@@

The code takes as input the size of the field, in terms of recoverable oil, and then if the field is smaller than 10 returns a year uniformly distributed between 1970 and 2013.  If the size is larger than 10, then the starting year is taken from a constrained normal distribution with a mean at 1975.  This is meant to reflect the common industry pattern of "creaming" - where large finds tend to be found earlier and which is apparent in the actual field-time distribution on the Norwegian Continental Shelf.

```{r}
gen_year<-function(size){
	#Generates initial year as a function of size/maxsize
	#let small fields be distributed uniformly from 1975 to 2013
	if(size<10){
		year<-trunc(runif(1, 1970, 2013))	
	} else{	
	#while large fields should be more common earlier - from a normal dist.
	#with mean at 1975
	#constrained to be between 1970 and 2013
		range<-FALSE
		while(range==FALSE){
			year<-trunc(rnorm(1,mean=(1975), sd=10))
			ifelse(year>=1970 & year<=2013, range<-TRUE, range<-FALSE)
			}
		}
return(year)
}	
```

@@

The below function encapsulates the two core operations involved in each step of the Monte Carlo experiment. The effect of price and uncertainty are added to each field via the "add_uncertainty()" function.  Then either a GAM or GLM regression is fun on the data.

@@

Below we plot the results in the form of emperical density functions of the results.  The vertical black lines represent the true values of beta.  Clearly, the GLM estimation has a significant negative bias.
```{r, fig.width=8, fig.height=5, message=FALSE}
ggplot(glm_mc_data, aes(x=coefficient, fill=beta)) +
geom_density(alpha=.5, size=0) +
geom_vline(aes(xintercept=c(0, 0.05))) +
scale_fill_grey() +
theme_bw() +
guides(fill=FALSE) +
xlab("beta_hat, estimated by GLM") +
ylab("Density")
```
### GAM Monte Carlo Experiment
Below we set the formula as our preffered GAM model with field-level random effects and smooth terms for both the production years and for calendar time, the latter restricted to have a 4-knot basis.

```{r}
formula_1<- formula(prod~s(prod_years, name, bs="re") + prices + size + s(year, bs="cr", k=4))
```

>>>>

First, we load some r packages that we will need:

@@

The function generates a starting year for field production from a half-normal distribution with a mean at 1970, but constrained to be between 1970 to 2000.  This is meant to reflect the common industry pattern that most significant finds are found early in a regions activity.

```{r}
gen_year<-function(size){
	#half-normal distribution with mean at 1970
	#constrained to be between 1970 and 2000
	range<-FALSE
	while(range==FALSE){
		year<-trunc(rnorm(1,mean=(1970), sd=10))
		ifelse(year>=1970 & year<=2000, range<-TRUE, range<-FALSE)
		}

return(year)
}	
```

@@

The below function encapsulates the two core operations involved in each step of the Monte Carlo experiment. The effect of price and uncertainty are added to each field via the "add_uncertainty()" function.  Then either a GAM or GLM regression is run on the data.

@@

Below we plot the results in the form of empirical density functions of the results.  The vertical black lines represent the true values of beta.  Clearly, the GLM estimation has a significant negative bias.
```{r, fig.width=8, fig.height=5, message=FALSE}
ggplot(glm_mc_data, aes(x=coefficient, fill=beta)) +
geom_density(alpha=.5, size=0) +
geom_vline(aes(xintercept=c(0, 0.05))) +
scale_fill_grey() +
theme_bw() +
guides(fill=FALSE) +
xlab("beta_hat, estimated by GLM") +
ylab("Density")
```
### GAM Monte Carlo Experiment
Below we set the formula as a slightly simplified version of our preffered GAM model. I have removed the cost index as this was not taken into account when generating the data.  For simplicity I also do not take into account lagged price terms as the main point of this monte-carlo experiment is to show how the production profile can be effectively controlled for.

```{r}
formula_1<- formula(prod~s(prod_years, size) + prices + year + I(year^2))
```
